---
title: "Associations between food category, food setting, hospitalization rate and death rate among salmonella outbreaks in the United States, 1998 to 2017"
output: html_notebook
---
################################################################################
###Preparing R Studio for negative bionomial regression
```{r}
# Load MASS for negative bin
library(MASS)
# Load ggplot for graphing
library(ggplot2)
# Load lmtest library for coeftest
library(lmtest)
## Loading required package: zoo
##
## Attaching package: 'zoo'
## The following objects are masked from 'package:base':
##
## as.Date, as.Date.numeric
# Load sandwich library for robust estimator
library(sandwich)
2
#load stargazer library to view a comparison of standard errors
library(stargazer)

install.packages("SMPracticals")
library(SMPracticals)
```

###Importing data
```{r}
MPH.5245.Final.project.data <- read.csv("~/Desktop/MPH 5245 Final project data/MPH 5245 Final project data.csv")

```


###First, analysing the association between hospitalizations and food category

###Preparing the dataset for the first analysis
```{r}
library(car)

#Recoding blanks as "NA"
MPH.5245.Final.project.data$IFSAC.Category[MPH.5245.Final.project.data$IFSAC.Category == ""]<-"NA"

#Recoding the IFSAC categories into simpler food groups
MPH.5245.Final.project.data$IFSAC.Category.Recoded<-recode(MPH.5245.Final.project.data$IFSAC.Category, "'Chicken'='Poultry'; 'Turkey'='Poultry'; 'Other Poultry'='Poultry'; 'Fish'='Seafood'; 'Crustaceans'='Seafood'; 'Fungi'='Vegetables & Related'; 'Grains-Beans'='Vegetables & Related'; 'Herbs'='Vegetables & Related'; 'Mollusks'='Seafood'; 'Nuts-Seeds'='Vegetables & Related'; 'Other Aquatic Animals'='Seafood'; 'Root/Underground'='Vegetables & Related'; 'Seeded Vegetables'='Vegetables & Related'; 'Vegetable Row Crops'='Vegetables & Related'; 'Game'='Other Meats'; 'Other Meat'='Other Meats'; 'Other'='Miscellaneous'; 'Unclassifiable'='Miscellaneous'; 'Multiple'='Miscellaneous'; 'Oils-Sugars'='Miscellaneous'")
```

###Preparing the offset
```{r}
#Collecting the relevant columns
PreliminaryAggregateFoodCategoryIllnessesHospitalizations<-aggregate(MPH.5245.Final.project.data, by=list(MPH.5245.Final.project.data$IFSAC.Category.Recoded, MPH.5245.Final.project.data$Illnesses, MPH.5245.Final.project.data$Hospitalizations), FUN=mean, na.rm=TRUE)

#Summing the number of illnesses by food category
AggregateFoodCategoryIllnesses<-aggregate(PreliminaryAggregateFoodCategoryIllnessesHospitalizations$Group.2,  by=list(Category=PreliminaryAggregateFoodCategoryIllnessesHospitalizations$Group.1), FUN=sum)

#Renaming the far-right column "Illnesses"
colnames(AggregateFoodCategoryIllnesses)[2] <- 'Illnesses'   

#Summing the number of hospitalizations by food category
AggregateFoodCategoryHospitalizations<-aggregate(PreliminaryAggregateFoodCategoryIllnessesHospitalizations$Group.3,  by=list(Category=PreliminaryAggregateFoodCategoryIllnessesHospitalizations$Group.1), FUN=sum)

#Renaming the far-right column "Hospitalizations"
colnames(AggregateFoodCategoryHospitalizations)[2] <- 'Hospitalizations' 

#Merging the three datasets together
FinalAggregateFoodCategoryIllnessesHospitalizations <- merge(AggregateFoodCategoryIllnesses, AggregateFoodCategoryHospitalizations, by="Category")
```

###Running the poisson regression between food category and hospitalization
```{r}
PoissonFoodCategoryHospitalization <- glm(Hospitalizations ~ Category + offset(log(Illnesses)), family = poisson(link = "log"), data = FinalAggregateFoodCategoryIllnessesHospitalizations)

#Presenting the results of the Poisson regression
summary(PoissonFoodCategoryHospitalization)
```

###Retrieving dispersion parameter from quasi-poisson
```{r}
QuasiPoissonFoodCategoryHospitalization <- glm(Hospitalizations ~ Category + offset(log(Illnesses)), family = quasipoisson(link = "log"), data = FinalAggregateFoodCategoryIllnessesHospitalizations)
summary(QuasiPoissonFoodCategoryHospitalization)
```

###Retrieving the rate ratios and 95% confidence intervals
```{r}
#Creating the function
glm.RR <- function(GLM.RESULT, digits = 2) {

    if (GLM.RESULT$family$family == "binomial") {
        LABEL <- "OR"
    } else if (GLM.RESULT$family$family == "poisson") {
        LABEL <- "RR"
    } else {
        stop("Not logistic or Poisson model")
    }

    COEF      <- stats::coef(GLM.RESULT)
    CONFINT   <- stats::confint(GLM.RESULT)
    TABLE     <- cbind(coef=COEF, CONFINT)
    TABLE.EXP <- round(exp(TABLE), digits)

    colnames(TABLE.EXP)[1] <- LABEL

    TABLE.EXP
}

#Exploiting the function
glm.RR(PoissonFoodCategoryHospitalization, 3)

#Results:
#
```

###Using robust standard errors through the robust sandwich covariance estimator and assigning it to an object to calculate IRRs
```{r}
RobustFoodCategoryHospitalization<-coeftest(PoissonFoodCategoryHospitalization, vcov = sandwich)
summary(RobustFoodCategoryHospitalization)
```


###Running the negative binomial regression
```{r}
NegativeBinomialFoodCategoryHospitalization<-glm.nb(Hospitalizations ~ Category + offset(log(Illnesses)), data = FinalAggregateFoodCategoryIllnessesHospitalizations)

library(lmtest)

lrtest(PoissonFoodCategoryHospitalization, NegativeBinomialFoodCategoryHospitalization)

#Negative binomial did not work. Therefore, Poisson with robust standard errors should be used.
```

###Calculating IRRs for Poisson + Robust Standard Errors model
```{r}
IRRFoodCategoryHospitalization <- cbind(IRR = coef(PoissonFoodCategoryHospitalization), "2.5%"=RobustFoodCategoryHospitalization[,1]-1.96*RobustFoodCategoryHospitalization[,2], "97.5%"=RobustFoodCategoryHospitalization[,1]+1.96*RobustFoodCategoryHospitalization[,2])
exp(IRRFoodCategoryHospitalization)

exp(coef(PoissonFoodCategoryHospitalization))
exp(confint(PoissonFoodCategoryHospitalization))

#The incidence rate of hospitalizations for -- 
# -- dairy-caused outbreaks is 1.30 times that of beef-related outbreaks. 
# -- fruit-caused outbreaks is 1.48 times that of beef-related outbreaks. 
# -- vegetable-caused outbreaks is 1.27 times that of beef-related outbreaks. 
```

################################################################################
################################################################################

###Second, let's analyse the association between deaths and food category

###Preparing the offset
```{r}
#Collecting the relevant columns
PreliminaryAggregateFoodCategoryIllnessesDeaths<-aggregate(MPH.5245.Final.project.data, by=list(MPH.5245.Final.project.data$IFSAC.Category.Recoded, MPH.5245.Final.project.data$Illnesses, MPH.5245.Final.project.data$Deaths), FUN=mean, na.rm=TRUE)

#Summing the number of deaths by food category
AggregateFoodCategoryDeaths<-aggregate(PreliminaryAggregateFoodCategoryIllnessesDeaths$Group.3,  by=list(Category=PreliminaryAggregateFoodCategoryIllnessesDeaths$Group.1), FUN=sum)

#Renaming the far-right column "Deaths"
colnames(AggregateFoodCategoryDeaths)[2] <- 'Deaths' 

#Merging the three datasets together
FinalAggregateFoodCategoryIllnessesDeaths <- merge(AggregateFoodCategoryIllnesses, AggregateFoodCategoryDeaths, by="Category")
```

###Running the poisson regression between food category and deaths
```{r}
PoissonFoodCategoryDeaths <- glm(Deaths ~ Category + offset(log(Illnesses)), family = poisson(link = "log"), data = FinalAggregateFoodCategoryIllnessesDeaths)

#Presenting the results of the Poisson regression
summary(PoissonFoodCategoryDeaths)
```

###Retrieving dispersion parameter from quasi-poisson
```{r}
QuasiPoissonFoodCategoryDeaths <- glm(Deaths ~ Category + offset(log(Illnesses)), family = quasipoisson(link = "log"), data = FinalAggregateFoodCategoryIllnessesDeaths)
summary(QuasiPoissonFoodCategoryDeaths)
```

###Retrieving the rate ratios and 95% confidence intervals
```{r}
glm.RR(PoissonFoodCategoryDeaths, 3)

#glm.RR did not work.

exp(coef(PoissonFoodCategoryDeaths))

#Not possible to find confidence intervals
```

###Using robust standard errors through the robust sandwich covariance estimator and assigning it to an object to calculate IRRs
```{r}
RobustFoodCategoryDeaths<-coeftest(PoissonFoodCategoryDeaths, vcov = sandwich)
summary(RobustFoodCategoryDeaths)
```

###Running the negative binomial regression
```{r}
NegativeBinomialFoodCategoryDeaths<-glm.nb(Deaths ~ Category + offset(log(Illnesses)), data = FinalAggregateFoodCategoryIllnessesDeaths)

lrtest(PoissonFoodCategoryDeaths, NegativeBinomialFoodCategoryDeaths)
#P-value > 0.05. P-value does not improve model fit. Therefore, Poisson model should be used. 
```

###Calculating IRRs for Poisson + Robust Standard Errors model
```{r}
IRRFoodCategoryDeaths <- cbind(IRR = coef(PoissonFoodCategoryDeaths), "2.5%"=RobustFoodCategoryDeaths[,1]-1.96*RobustFoodCategoryDeaths[,2], "97.5%"=RobustFoodCategoryDeaths[,1]+1.96*RobustFoodCategoryDeaths[,2])
exp(IRRFoodCategoryDeaths)

exp(coef(PoissonFoodCategoryDeaths))
exp(confint(PoissonFoodCategoryDeaths))

#The incidence rate of deaths for -- 
# -- fruit-caused outbreaks is 4.88 times that of beef-related outbreaks. 
# -- vegetable-caused outbreaks is 2.63 times that of beef-related outbreaks. 
# -- sprouts-caused outbreaks is 1.31 times that of beef-related outbreaks. 
# -- seafood-caused outbreaks is 0.00 times that of beef-related outbreaks. 

#Cannot find confidence intervals
```

################################################################################
################################################################################
################################################################################

###Third, let's analyse the association between hospitalization and food setting

###Recoding settings into simpler categories based on which one is dominant or 
###comes first
```{r}
library(dplyr)

MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Banquet Facility (food prepared and served on-site)"] <- "Other"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Banquet Facility (food prepared and served on-site)"] <- "Other"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Banquet Facility (food prepared and served on-site); Caterer (food prepared off-site from where served)"] <- "Other"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Banquet Facility (food prepared and served on-site); Caterer (food prepared off-site from where served); Religious facility"] <- "Other"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Banquet Facility (food prepared and served on-site); Other"] <- "Other"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Camp"] <- "Other"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Camp; Other"] <- "Other"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Caterer (food prepared off-site from where served)"] <- "Catering"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Caterer (food prepared off-site from where served); Child day care"] <- "Catering"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Caterer (food prepared off-site from where served); Fair, festival, other temp or mobile services"] <- "Catering"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Caterer (food prepared off-site from where served); Fair, festival, other temp or mobile services; Grocery store"] <- "Catering"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Caterer (food prepared off-site from where served); Grocery store"] <- "Catering"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Caterer (food prepared off-site from where served); Grocery store; Other"] <- "Catering"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Caterer (food prepared off-site from where served); Other"] <- "Catering"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Caterer (food prepared off-site from where served); School/college/university"] <- "Catering"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Child day care"] <- "Educational Facility"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Fair, festival, other temp or mobile services"] <- "Mobile"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Fair, festival, other temp or mobile services; Grocery store"] <- "Mobile"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Farm/dairy"] <- "Other"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Grocery store"] <- "Retail"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Grocery store; Office/indoor workplace"] <- "Retail"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Grocery store; Other"] <- "Retail"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Hospital"] <- "Health Care Facility"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Long-term care/nursing home/assisted living facility"] <- "Health Care Facility"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Long-term care/nursing home/assisted living facility; Hospital"] <- "Health Care Facility"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Long-term care/nursing home/assisted living facility; Other"] <- "Health Care Facility"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Office/indoor workplace"] <- "Other"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Office/indoor workplace; Hospital"] <- "Other"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Office/indoor workplace; Other"] <- "Other"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Other"] <- "Other"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Other; Unknown"] <- "Other"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Prison/jail"] <- "Other"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence; Banquet Facility (food prepared and served on-site)"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence; Banquet Facility (food prepared and served on-site); Caterer (food prepared off-site from where served)"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence; Banquet Facility (food prepared and served on-site); Other"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence; Caterer (food prepared off-site from where served)"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence; Caterer (food prepared off-site from where served); Grocery store"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence; Caterer (food prepared off-site from where served); Other"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence; Child day care"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence; Fair, festival, other temp or mobile services"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence; Fair, festival, other temp or mobile services; Grocery store"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence; Fair, festival, other temp or mobile services; Other"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence; Grocery store"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence; Grocery store; Child day care"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence; Grocery store; Long-term care/nursing home/assisted living facility"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence; Grocery store; Long-term care/nursing home/assisted living facility; School/college/university"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence; Grocery store; Other"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence; Grocery store; Religious facility"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence; Grocery store; Religious facility; Other"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence; Long-term care/nursing home/assisted living facility"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence; Office/indoor workplace"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence; Other"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence; Religious facility"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Private home/residence; School/college/university"] <- "Residence"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Religious facility"] <- "Other"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Fast-food(drive up service or pay at counter)"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Fast-food(drive up service or pay at counter); Grocery store"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Fast-food(drive up service or pay at counter); Private home/residence"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Fast-food(drive up service or pay at counter); Private home/residence; Grocery store"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Fast-food(drive up service or pay at counter); Restaurant - Sit-down dining"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Fast-food(drive up service or pay at counter); Restaurant - Sit-down dining; Fair, festival, other temp or mobile services"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Fast-food(drive up service or pay at counter); Restaurant - Sit-down dining; Office/indoor workplace; Other"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Fast-food(drive up service or pay at counter); Restaurant - Sit-down dining; Private home/residence"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Fast-food(drive up service or pay at counter); Restaurant - Sit-down dining; Restaurant - other or unknown type"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Fast-food(drive up service or pay at counter); Restaurant - Sit-down dining; Restaurant - other or unknown type; Grocery store; Office/indoor workplace"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Fast-food(drive up service or pay at counter); Restaurant - Sit-down dining; Restaurant - other or unknown type; Private home/residence"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Fast-food(drive up service or pay at counter); Restaurant - Sit-down dining; Restaurant - other or unknown type; Private home/residence; Grocery store; Long-term care/nursing home/assisted living facility; Camp"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Fast-food(drive up service or pay at counter); Restaurant - Sit-down dining='Restaurant'; School/college/university"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Buffet"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Banquet Facility (food prepared and served on-site)"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Caterer (food prepared off-site from where served)"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Caterer (food prepared off-site from where served); Fair, festival, other temp or mobile services"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Fair, festival, other temp or mobile services"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Grocery store"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Grocery store; Other"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Grocery store; School/college/university"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Hospital; School/college/university"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Long-term care/nursing home/assisted living facility"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Long-term care/nursing home/assisted living facility; Hospital"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Long-term care/nursing home/assisted living facility; Other"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Office/indoor workplace"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Office/indoor workplace; Hospital"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Other"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Prison/jail"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Private home/residence"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Private home/residence; Banquet Facility (food prepared and served on-site); Grocery store"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Private home/residence; Caterer (food prepared off-site from where served)"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Private home/residence; Caterer (food prepared off-site from where served); Grocery store; Long-term care/nursing home/assisted living facility; Hospital"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Private home/residence; Fair, festival, other temp or mobile services"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Private home/residence; Fair, festival, other temp or mobile services"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Private home/residence; Grocery store"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Private home/residence; Grocery store; Long-term care/nursing home/assisted living facility; Other"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Private home/residence; Grocery store; Other"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Private home/residence; Grocery store; Unknown"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Private home/residence; Hospital"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Private home/residence; Other"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Religious facility"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; School/college/university"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - other or unknown type; Unknown"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Sit-down dining"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Sit-down dining; Banquet Facility (food prepared and served on-site)"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Sit-down dining; Caterer (food prepared off-site from where served)"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Sit-down dining; Caterer (food prepared off-site from where served); Religious facility"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Sit-down dining; Fair, festival, other temp or mobile services"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Sit-down dining; Grocery store"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Sit-down dining; Long-term care/nursing home/assisted living facility; Unknown"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Sit-down dining; Office/indoor workplace; School/college/university"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Sit-down dining; Other"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Sit-down dining; Private home/residence"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Sit-down dining; Private home/residence; Banquet Facility (food prepared and served on-site); Caterer (food prepared off-site from where served); Grocery store"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Sit-down dining; Private home/residence; Grocery store"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Sit-down dining; Private home/residence; Other"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Sit-down dining; Restaurant - Buffet"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Sit-down dining; Restaurant - other or unknown type; Private home/residence"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Sit-down dining; Restaurant - other or unknown type; Private home/residence; Banquet Facility (food prepared and served on-site); Caterer (food prepared off-site from where served)"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Restaurant - Sit-down dining; Unknown"] <- "Restaurant"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="School/college/university"] <- "Educational Facility"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="School/college/university; Other"] <- "Educational Facility"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting=="Unknown"] <- "Other"
MPH.5245.Final.project.data$Setting.Recoded[MPH.5245.Final.project.data$Setting==""] <- "NA"
```

###Preparing the data
```{r}
#Collecting the relevant columns
PreliminaryAggregateSettingIllnessesHospitalizations<-aggregate(MPH.5245.Final.project.data, by=list(MPH.5245.Final.project.data$Setting.Recoded, MPH.5245.Final.project.data$Illnesses, MPH.5245.Final.project.data$Hospitalizations), FUN=mean, na.rm=TRUE)

#Summing the number of illnesses by setting
AggregateSettingIllnesses<-aggregate(PreliminaryAggregateSettingIllnessesHospitalizations$Group.2,  by=list(Category=PreliminaryAggregateSettingIllnessesHospitalizations$Group.1), FUN=sum)

#Renaming the far-right column "Illnesses"
colnames(AggregateSettingIllnesses)[2] <- 'Illnesses'   

#Summing the number of hospitalizations by food category
AggregateSettingHospitalizations<-aggregate(PreliminaryAggregateSettingIllnessesHospitalizations$Group.3,  by=list(Category=PreliminaryAggregateSettingIllnessesHospitalizations$Group.1), FUN=sum)

#Renaming the far-right column "Hospitalizations"
colnames(AggregateSettingHospitalizations)[2] <- 'Hospitalizations' 

#Merging the three datasets together
FinalAggregateSettingIllnessesHospitalizations <- merge(AggregateSettingIllnesses, AggregateSettingHospitalizations, by="Category")

#Renaming the far-left column "Setting"
colnames(FinalAggregateSettingIllnessesHospitalizations)[1] <- 'Setting' 
```

###Running the poisson regression between setting and hospitalization
```{r}
#Running the poisson regression
PoissonSettingHospitalization <- glm(Hospitalizations ~ Setting + offset(log(Illnesses)), family = poisson(link = "log"), data = FinalAggregateSettingIllnessesHospitalizations)

#Presenting the results of the Poisson regression
summary(PoissonSettingHospitalization)
```

###Retrieving dispersion parameter from quasi-poisson
```{r}
QuasiPoissonSettingHospitalization <- glm(Hospitalizations ~ Setting + offset(log(Illnesses)), family = quasipoisson(link = "log"), data = FinalAggregateSettingIllnessesHospitalizations)
summary(QuasiPoissonSettingHospitalization)
```

###Retrieving the rate ratios and 95% confidence intervals
```{r}
#Exploiting the function
glm.RR(PoissonSettingHospitalization, 3)
```

###Using robust standard errors through the robust sandwich covariance estimator and assigning it to an object to calculate IRRs
```{r}
RobustSettingHospitalization<-coeftest(PoissonSettingHospitalization, vcov = sandwich)
summary(RobustSettingHospitalization)
```

###Running the negative binomial regression
```{r}
NegativeBinomialSettingHospitalization<-glm.nb(Hospitalizations ~ Setting + offset(log(Illnesses)), data = FinalAggregateSettingIllnessesHospitalizations)

library(lmtest)

lrtest(PoissonSettingHospitalization, NegativeBinomialSettingHospitalization)

#P<0.05. P-value improves model fit. Therefore, the negative binomial model should be used.
```

###Calculating IRRs for Poisson + Robust Standard Errors model
```{r}
IRRSettingHospitalization <- cbind(IRR = coef(NegativeBinomialSettingHospitalization), "2.5%"=RobustSettingHospitalization[,1]-1.96*RobustSettingHospitalization[,2], "97.5%"=RobustSettingHospitalization[,1]+1.96*RobustSettingHospitalization[,2])
exp(IRRSettingHospitalization)

exp(coef(NegativeBinomialSettingHospitalization))
exp(confint(NegativeBinomialSettingHospitalization))

#The incidence rate of hospitalizations for -- 
# -- outbreaks occurring in healthcare facilities is 2.37 times that of outbreaks occurring from catering. 
# -- outbreaks occurring in residences is 2.44 times that of outbreaks occurring from catering.
# -- outbreaks occurring in restaurants is 1.77 times that of outbreaks occurring from catering. 
```

################################################################################
################################################################################
################################################################################
################################################################################

###Finally, let's analyse the association between deaths and food setting

###Preparing the offset
```{r}
#Collecting the relevant columns
PreliminaryAggregateSettingIllnessesDeaths<-aggregate(MPH.5245.Final.project.data, by=list(MPH.5245.Final.project.data$Setting.Recoded, MPH.5245.Final.project.data$Illnesses, MPH.5245.Final.project.data$Deaths), FUN=mean, na.rm=TRUE)

#Summing the number of deaths by food category
AggregateSettingDeaths<-aggregate(PreliminaryAggregateSettingIllnessesDeaths$Group.3,  by=list(Category=PreliminaryAggregateSettingIllnessesDeaths$Group.1), FUN=sum)

#Renaming the far-right column "Deaths"
colnames(AggregateSettingDeaths)[2] <- 'Deaths' 

#Merging the three datasets together
FinalAggregateSettingIllnessesDeaths <- merge(AggregateSettingIllnesses, AggregateSettingDeaths, by="Category")

#Renaming the far-left column "Setting"
colnames(FinalAggregateSettingIllnessesDeaths)[1] <- 'Setting' 
```

###Running the poisson regression between setting and death
```{r}
#Running the poisson regression
PoissonSettingDeath <- glm(Deaths ~ Setting + offset(log(Illnesses)), family = poisson(link = "log"), data = FinalAggregateSettingIllnessesDeaths)

#Presenting the results of the Poisson regression
summary(PoissonSettingDeath)
```

###Retrieving dispersion parameter from quasi-poisson
```{r}
QuasiPoissonSettingDeath <- glm(Deaths ~ Setting + offset(log(Illnesses)), family = quasipoisson(link = "log"), data = FinalAggregateSettingIllnessesDeaths)
summary(QuasiPoissonSettingDeath)
```

###Retrieving the rate ratios and 95% confidence intervals
```{r}
#Exploiting the function
glm.RR(PoissonSettingDeath, 3)

exp(coef(PoissonSettingDeath))

exp(confint(PoissonSettingDeath))

#Confidence intervals cannot be found
```

###Using robust standard errors through the robust sandwich covariance estimator and assigning it to an object to calculate IRRs
```{r}
RobustSettingDeath<-coeftest(PoissonSettingDeath, vcov = sandwich)
summary(RobustSettingDeath)
```

###Running the negative binomial regression
```{r}
NegativeBinomialSettingDeath<-glm.nb(Deaths ~ Setting + offset(log(Illnesses)), data = FinalAggregateSettingIllnessesDeaths)

library(lmtest)

lrtest(PoissonSettingDeath, NegativeBinomialSettingDeath)

#P>0.05. P-value does not improve model fit. Therefore, the poisson model should be used.
```

###Calculating IRRs for Poisson + Robust Standard Errors model
```{r}
IRRSettingDeath <- cbind(IRR = coef(PoissonSettingDeath), "2.5%"=RobustSettingDeath[,1]-1.96*RobustSettingDeath[,2], "97.5%"=RobustSettingDeath[,1]+1.96*RobustSettingDeath[,2])
exp(IRRSettingDeath)

exp(coef(PoissonSettingDeath))
exp(confint(PoissonSettingDeath))

#The incidence rate of death for -- 
# -- outbreaks occurring in healthcare facilities is 40.84 times that of outbreaks occurring from catering. 
# -- outbreaks occurring in residences is 4.72 times that of outbreaks occurring from catering.
# -- outbreaks occurring in retail is 4.78 times that of outbreaks occurring from catering.
# -- outbreaks occurring in restaurants is 1.43 times that of outbreaks occurring from catering. 
```
